\documentclass{math}

\title{Introduction to Computer Science Theory}
\author{Alvin Lin}
\date{August 2017 - December 2017}

\begin{document}

\maketitle

\section*{Pumping Lemmas}
A general criteria for establishing non-regularity: If \( A \) is a regular
language, then there is an integer \( p \) such that for any \( s\in A \) with
\( |s|\ge p \) there are strings \( x,y,z \) such \( s = xyz \) and:
\begin{enumerate}
  \item for any \( i\ge0, xy^iz\in A \)
  \item \( |y| > 0 \)
  \item \( |xy|\le p \)
\end{enumerate}
This is called a pumping lemma because you repeat or pump the substring \( y \).

\subsection*{Using the pumping lemma}
The language \( L = \{a^jb^{2j}\mid j\ge 0\} \) is not regular. To prove this,
suppose for a contradiction that \( L \) is regular. Then by the pumping lemma,
let \( p \) be the integer such that every string \( s \) in \( L \) with
\( |s|\ge p \) can be written as \( xyz \), where \( |xy|\le p, |y| > 0 \), and
\( xy^iz\in L \) for all \( i\ge0 \).
\begin{itemize}
  \item Let \( s = a^pb^{2p} \). Then \( s\in L \) and \( |s|\ge p \). So, there
  exist strings \( x,y,z \) such that \( s = xyz, |xy|\le p, |y|>0 \), and
  \( xy^iz\in L \) for all \( i\ge0 \).
  \item In particular, \( xz\in L \) (choosing \( i = 0 \))
  \item Since \( |xy|\le p \) and \( |y| > 0 \), \( y \) consists of 1 or more
  \( a \)'s. If we remove \( y \) from \( a^pb^{2p} \), we get a string with
  \( 2p~b \)'s and less than \( p~a \)'s.
  \item \( xz \) has \( 2p~b \)'s and less than \( p~a \)'s. But then, by the
  definition of \( L \), \( xz\notin L \). This is a contradiction.
  \item It follows that the assumption that \( L \) is regular is wrong.
  \item So, we have shown that \( L \) is not regular.
\end{itemize}

\subsubsection*{Example}
Prove \( L \) is not regular.
\[ L = \{a^{p^2}\mid p\ge0\} \]
Assume that \( L \) is regular. By the pumping lemma, there is a \( p\in\N \)
such that for all \( s\in L \) such that \( |s|\ge p \) there exists \( x,y,z \)
such that \( s = xyz, |xy|\le p, |y|>0 \) for all \( i\ge0, xy^iz\in L \). Let
\( s = a^{p^2} \). So by definition of \( L \), \( s\in L \) and \( |s| \) is
clearly greater than or equal to \( p \). So, there exist \( x,y,z \) such that
\( s = xyz, |xy|\le p, |y|>0 \) and for all \( i\ge0, xy^iz\in L \). Let
\( i = 2 \). \par
By the pumping lemma, \( xy^2z\in L \). Since \( |y|>0,|xy^2z|>p^2 \). Thus
\( |xy^2z|\ge(p+1)^2 \). Since \( |xy|\le p, |xy^2z|\le p^2+p \). \par
So, \( (p+1)^2 \le p^2+p \). This is a contradiction and thus \( L \) is not
regular.

\subsubsection*{Example}
\[ \Sigma = \left\{
  \begin{bmatrix}0 \\ 0\end{bmatrix},
  \begin{bmatrix}0 \\ 1\end{bmatrix},
  \begin{bmatrix}1 \\ 0\end{bmatrix},
  \begin{bmatrix}1 \\ 1\end{bmatrix}
\right\} \]
\[ L = \{x\in\Sigma^*\mid\text{the top string is the reverse of the bottom}\} \]
\[ \begin{bmatrix}0 \\ 1\end{bmatrix}
  \begin{bmatrix}1 \\ 1\end{bmatrix}
  \begin{bmatrix}1 \\ 0\end{bmatrix}\in L \]
\[ \begin{bmatrix}0 \\ 0\end{bmatrix}
  \begin{bmatrix}1 \\ 1\end{bmatrix}
  \begin{bmatrix}0 \\ 0\end{bmatrix}\in L \]
\[ \begin{bmatrix}0 \\ 1\end{bmatrix}
  \begin{bmatrix}0 \\ 1\end{bmatrix}
  \begin{bmatrix}1 \\ 0\end{bmatrix}\notin L \]
\[ \begin{bmatrix}0 \\ 0\end{bmatrix}
  \begin{bmatrix}1 \\ 0\end{bmatrix}\notin L \]
Assume that \( L \) is regular. By the pumping lemma, there exists \( p\in\N \)
such that for all \( s\in L \) such that \( |s|\ge p \) there exist \( x,y,z \)
such that \( s = xyz, |xy|\le p, |y|>0 \), and \( \forall{i}\ge0(xy^iz\le L) \).
\par
Let \( s = \begin{bmatrix}0 \\ 1\end{bmatrix}^p
\begin{bmatrix}1 \\ 0\end{bmatrix}^p \). Let \( x,y,z \) be such that
\( s = xyz, |xy|\le p, |y|>0 \) and for all \( i\ge0, xy^iz\in L \). \par
Let \( i = 3 \). By the pumping lemma, \( xy^3z\in L \). But \( xy^3z =
\begin{bmatrix}0 \\ 1\end{bmatrix}^{p+2|y|}
\begin{bmatrix}1 \\ 0\end{bmatrix}^p \). This is a contradiction and thus
\( L \) is not regular.

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, please contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
