\documentclass{math}

\title{Boundary Value Problems}
\author{Alvin Lin}
\date{August 2018 - December 2018}

\begin{document}

\maketitle

\section*{Introduction and Classifications}
Recall that ordinary differential equations (ODEs) were often focused on
initial value problems. For example:
\[ y''+3y'+2y = 0 \quad y(x_0) = 3 \quad y'(x_0) = -1 \]
Boundary value problems consist of a differential equation and boundary
conditions given at two or more values of the independent variable. For example:
\[ y''+16y = 0 \quad y(x_0) = y_0 \quad y(x_1) = y_1 \]
It is implied here that \( x_0 \ne x_1 \). Boundary conditions can vary greatly:
\begin{align*}
  y'(x_0) &= y_0 \quad y(x_1) = y_1 \\
  y(x_0) &= y_0 \quad y'(x_1) = y_1 \\
  y'(x_0) &= y_0 \quad y'(x_1) = y_1
\end{align*}
Notationally, we may write \( \ddiff{y}{x} \) as \( D_y \) where \( D \) is an
operator applied to the variable \( y \). Similarly, we may also abbreviate
\( \ddiff{^2y}{x^2} \) as \( DDy \) or \( D^2y \). For example, the following
two differential equations are equivalent:
\begin{align*}
  y''+3y'+2y &= 0 \\
  D^2y+3Dy+2y &= 0
\end{align*}
Suppose we define the operator \( L = D^2+3D+2 \) (short for ``linear''
operator), then \( Ly = 0 \). Linear operations on \( y \) include
\( y, y', y'', 3y, 3y', 3xy, 5xy' \), etc. Non-linear operations on \( y \)
include \( y^2, \sin(y), yy' \), etc. Note that \( L \) is a linear operator
if and only if \( L(c_1y_1+c_2y_2) = c_1Ly_1+c_2Ly_2 \). For example:
\begin{itemize}
  \item \( y'+2xy = 0 \) is a linear differential equation.
  \item \( y'+3x\sin(y) = 0 \) is a non-linear differential equation.
\end{itemize}
For an ordinary differential equation, the order of the equation is the highest
order of derivative that appears in the equation. Partial differential equations
may contain partial derivatives, and their order is also defined by the highest
partial derivative it contains. \par
The generic form of a linear differential operator is as follows:
\[ L = \alpha_n(x)D^n+\alpha_{n-1}D^{n-1}+\dots+\alpha_1(x)D+\alpha_0(x) \]
where \( \alpha \) can either be a constant or coefficient in terms of \( x \).
If \( \alpha_0,\alpha_1,\dots,\alpha_n \) are all constants, then the
differential equation is a constant coefficient equation. If any \( \alpha \)
depends explicitly on \( x \), then the differential equation is a variable
coefficient equation. \par
If an equation has the form \( Ly = 0 \), then the differential equation is
linear and homogeneous. If an equation has the form \( L(y) = f(x) \ne 0 \),
then we call the equation non-homogeneous. These properties all influence the
type of solution we will use as well as the ease of solving the differential
equation.
\begin{center}
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    linear & low order & homogeneous & const coefficient & ODE \\
    \hline
    non-linear & high order & non-homogeneous & variable coefficient & PDE \\
    \hline
  \end{tabular}
\end{center}
\begin{itemize}
  \item \( 2\sin(x)y''+y+6 = 0 \) \\
    ODE \\
    linear \\
    second order \\
    variable coefficient \\
    non-homogeneous
  \item \( uu_{xxx}+xu_t = 0 \) \\
    PDE \\
    non-linear \\
    third order
\end{itemize}

\subsection*{Standard Form of a Second Order Linear PDE}
A PDE is linear and second order if it can be written in the form:
\[ A\pdiff{^2u}{x^2}+B\pdiff{^2u}{x\partial y}+C\pdiff{^2u}{y^2}+
  D\pdiff{u}{x}+E\pdiff{u}{y}+Fu = G \]
with addition and subtraction of terms on both sides where \( u \) is the
dependent variable, \( x,y \) are the independent variables, and
\( A,B,C,D,E,F,G \) are constants or explicit functions of \( x \) or \( y \).
\par
If \( G(x,y) \equiv 0 \)), then the PDE is homogeneous, otherwise it is
non-homogeneous. The PDE is:
\begin{itemize}
  \item hyperbolic if \( B^2-4AC > 0 \)
  \item parabolic if \( B^2-4AC = 0 \)
  \item elliptic if \( B^2-4AC < 0 \)
\end{itemize}

\subsubsection*{Example: Heat Equation}
\( u(x,t) \) represents the temperature along a rod at time \( t \) and
position \( x \).
\[ u_t = \beta u_{xx} \]
where \( \beta \) is real, constant, and greater than 0. If we rewrite this
according to our standard form:
\begin{align*}
  \beta u_{xx}+0u_{xt}+0u_{tt}+0u_x-u_t+0 &= 0 \\
  \beta u_{xx}-u_t &= 0 \\
  A &= \beta \quad B = 0 \quad C = 0 \\
  B^2-4AC &= 0
\end{align*}
Therefore, this equation is parabolic.

\subsection*{Solving PDEs}
Throughout this course, we will build a procedure for solving PDEs:
\begin{itemize}
  \item Use separation of variables if possible to break a PDE into two more
    ODEs.
  \item Solve each ODE subproblem.
  \item Assemble a general solution out of the ODE solutions.
  \item Solve for coefficients in the general solution using the boundary
    conditions or initial conditions if available.
\end{itemize}

\subsubsection*{Example}
\begin{align*}
  y''+16y &= 0 \quad y(0) = 0 \quad y(\frac{\pi}{2}) = 0 \\
  y &= \e^{rt} \\
  r^2+16 &= 0 \\
  r &= 0\pm4i \quad \alpha = 0 \quad \beta = 4 \\
  y &= \e^{\alpha t}(c_1\cos(\beta t)+c_2\sin(\beta t)) \\
  &= c_1\cos(4x)+c_2\sin(4x)
\end{align*}
This is a general solution to the ODE. We can apply the boundary conditions to
try and solve for the unknown coefficients.
\begin{align*}
  y(0) &= 0 \\
  &= c_1\cos(0)+c_2\sin(0) \\
  c_1 &= 0 \\
  y(\frac{\pi}{2}) &= 0 \\
  &= c_1\cos\left(\frac{4\pi}{2}\right)+c_2\sin\left(\frac{4\pi}{2}\right) \\
  &= c_2\sin\left(\frac{4\pi}{2}\right) \\
  &= 0
\end{align*}
We can determine a specific answer for \( c_1 \), but \( c_2 \) can be any real
number.
\[ y = c_2\sin(4x) \quad c_2\in\R \]
If we were to change the boundary conditions to \( y(0) = 0 \) and
\( y(\frac{\pi}{2}) = 1 \), then the boundary value problem would have no
solution. Boundary value problems can have no solution, exactly one solution,
or an infinite number of solutions.

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, please contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
