\documentclass{math}

\usepackage{tikz}

\title{Boundary Value Problems}
\author{Alvin Lin}
\date{August 2018 - December 2018}

\begin{document}

\maketitle

\section*{Separation of Variables}
Knowing if a partial differential equation is separable can help us solve it.
Suppose we have the function \( u_t = 8u_{xx} \). To test for separability:
\begin{enumerate}
  \item Suppose \( u(x,t) = X(x)T(t) \).
  \item We need to substitute this into the partial differential equation.
  \item If you can isolate all instances of the \( x \) independent variable
    on one side and the \( t \) independent variable on the other side of the
    equation, then the partial differential equation is separable.
\end{enumerate}
\begin{align*}
  u(x,t) &= X(x)T(t) \\
  \pdiff{}{t}(X(x)T(t)) &= 8\pdiff{^2}{x^2}(X(x)T(t)) \\
  X(x)\pdiff{}{t}T(t) &= 8\pdiff{^2}{x^2}(X(x))T(t) \\
  X(x)T'(t) &= 8X''(x)T(t) \\
  \frac{X(x)}{X''(x)} &= \frac{8T'(t)}{T(t)}
\end{align*}
In their separated form, we can set both sides equal to some separation constant
to solve two ordinary differential equations. This separation constant is
usually denoted \( \lambda, -\lambda, \alpha^2, \text{ or } -\alpha^2 \).
\[ \frac{X(x)}{X''(x)} = \frac{8T'(t)}{T(t)} = -\lambda \]
Separation is useful because we can break a partial differential equation into
multiple ordinary differential equations.

\subsubsection*{Example}
Generic form of the heat equation:
\[ \pdiff{u}{t} = \beta\pdiff{^2u}{x^2} \]
We want to find a solution that is nontrivial and bounded. It helps to narrow
things down if we have boundary conditions and/or initial conditions.
\[ u(0,t) = 0 \quad u(L,t) = 0 \quad u(x,0) = f(x) \]
We want to find \( u(x,t) \) that satisfies the partial differential equation,
its boundary conditions, and its initial condition. Since the equation is
separable, we can substitute \( u(x,t) = X(x)T(t) \) into the equation and show
that
\[ \frac{X''(x)}{X(x)} = \frac{T'(t)}{\beta T(t)} \]
If we choose the separation constant \( -\lambda \):
\begin{align*}
  \frac{X''(x)}{X(x)} &= -\lambda = \frac{T'(t)}{\beta T(t)} \\
  X''(x) &= -\lambda X(x) \\
  T'(t) &= -\lambda\beta T(t)
\end{align*}
With the boundary condition \( u(0,t) = 0 \), we know that \( X(0)T(t) = 0 \).
With the boundary condition \( u(L,t) = 0 \), we know that \( X(L)T(t) = 0 \).
\begin{align*}
  u(x,0) &= X(x)T(0) = f(x) \\
  X(x) &= \frac{f(x)}{T(0)}
\end{align*}
We can start with either ordinary differential equation:
\begin{align*}
  X''(x) &= -\lambda X(x) \quad X(0) = 0 \quad X(L) = 0 \\
  X''(x)+\lambda X(x) &= 0 \\
  r^2+\lambda &= 0
\end{align*}
Using this characteristic equation, we have three cases:
\begin{enumerate}
  \item \( \lambda < 0 \)
  \begin{align*}
    r^2 &= -\lambda \\
    r_{1,2} &= \pm\sqrt{-\lambda} \\
    X(x) &= c_1\e^{\sqrt{-\lambda}x}+c_2\e^{-\sqrt{\lambda}x} \\
    X(0) &= 0 = c_1\e^{0}+c_2\e^{0} \\
    c_1 &= -c_2 \\
    X(L) &= 0 = c_1(\e^{\sqrt{-\lambda}L}-\e^{-\sqrt{\lambda}L}) \\
    0 &= c_1(\e^{2\sqrt{-\lambda}L}-1) \\
    c_1 &= 0 = c_2 \\
    X(x) &= 0 \quad \text{(trivial solution)}
  \end{align*}
  \item \( \lambda = 0 \)
  \begin{align*}
    X''(x) &= 0 \\
    r^2 &= 0 \\
    r &= 0 \\
    X(x) &= c_1\e^{0t}+c_2x\e^{0t} \\
    &= c_1+xc_2 \\
    X(0) &= 0 = c_1 \\
    c_1 &= 0 \\
    X(L) &= xc_2 \\
    c_2 &= 0 \\
    X(x) &= 0 \quad \text{(trivial solution)}
  \end{align*}
  \item \( \lambda > 0 \)
  \begin{align*}
    r^2+\lambda &= 0 \\
    r_{1,2} &= \pm\sqrt{-\lambda} \\
    \alpha^2 &= -\lambda \\
    r_{1,2} &= \pm i\alpha = \pm i\sqrt{\lambda} \\
    X(x) &= c_1\cos(\alpha x)+c_2\sin(\alpha x) \\
    X(0) &= 0 = c_1\cos(0)+c_2\sin(0) \\
    c_1 &= 0 \\
    X(L) &= 0 = c_2\sin(\alpha L)
  \end{align*}
  If \( c_2 \) is 0, then the solution is again trivial. If it is not zero,
  then the solution depends on the separation constant.
  \begin{align*}
    \sin(\alpha L) &= 0 \\
    \alpha L &= n\pi, \quad n\in\Z \\
    \alpha &= \frac{n\pi}{L} \\
    X(x) &= c_2\sin\left(\frac{n\pi}{L}x\right)
  \end{align*}
\end{enumerate}
To produce a non-trivial solution for \( X(x) \), we needed to use
\( \sqrt{\lambda} = \frac{n\pi}{L} \). We can use this to solve the second
ordinary differential equation.
\begin{align*}
  T'(t) &= -\lambda\beta T(t) \\
  T'(t) &= -\frac{n^2\pi^2}{L^2}\beta T(t) \\
  T(t) &= c_2\e^{ct} \\
  T'(t) &= \ddiff{}{t}(T(t)) \\
  &= c_2c\e^{ct} \\
  &= cT(t) \\
  T(t) &= c_3\e^{-\frac{n^2\pi^2}{L^2}\beta t}
\end{align*}
We can substitute this back into \( u(x,t) = X(x)T(t) \) to find our solution.
\begin{align*}
  u(x,t) &= X(x)T(t) \\
  &= c_2\sin(\frac{n\pi x}{L})c_2\e^{-\frac{n^2\pi^2}{L^2}\beta t} \\
  u_n(x,t) &= c_n\sin(\frac{n\pi x}{L})\e^{-(\frac{n\pi}{L})^2\beta t}
\end{align*}
This has infinitely many solutions because haven't applied our initial condition
yet.

\subsection*{Eigenvalues and Eigenfunctions}
Thus far, we have found special \( X(x) \) and \( \lambda \) values that made
the boundary value problem solutions non-trivial. These are known as
eigenfunctions:
\begin{align*}
  X_n(x) &= c_n\sin(\frac{n\pi x}{L}) = \text{eigenfunctions of the BVP} \\
  \lambda_n &= (\frac{n\pi}{L})^2 = \text{eigenvalues of the BVP}
\end{align*}
We can associate eigenvalues and eigenfunctions with operators.
\begin{align*}
  X''+\lambda X &= 0 \\
  X'' &= -\lambda X \\
  LX &= -\lambda X \\
  Ly &= D^2y = y'' \\
  LX &= D^2X = X''
\end{align*}
Eigenvalues of a linear operator \( L \) are values for which \( Ly =
\lambda y \) has non-trivial solutions for the given boundary conditions.
Eigenfunctions are functions \( y_n(x) \) corresponding to eigenvalues
\( \lambda_n \). Problems that require you to find the eigenvalues and
eigenfunctions of boundary value problems require you to find all non-trivial
\( y_n \) solutions.

\subsubsection*{Superposition Principle}
If \( u_1,u_2,\dots,u_n \) are solutions to a linear homogeneous differential
equation, then \( u(x,t) = \sum_{i=1}^{\infty}c_iu_i(x,t) \) is also a solution
where \( c_i \) are arbitrary constants. If we look at the solution we acquired
before:
\[ u_n(x,t) = c_n\sin(\frac{n\pi x}{L})\e^{-(\frac{n\pi}{L})^2\beta t} \]
We can rewrite this as:
\[ u(x,t) = \sum_{n=1}^{\infty}c_n\sin(\frac{n\pi x}{L})
  \e^{-(\frac{n\pi}{L})^2\beta t} \]

\subsection*{Summary}
To recap, these are the steps to solve this type of boundary value problem:
\begin{enumerate}
  \item Separate the variables.
  \[ u(x,t) = X(x)T(t) \]
  Plug these back into the partial differential equation to get two ordinary
  differential equations related by a separation constant \( \lambda \) or
  \( \alpha^2 \).
  \item Use the boundary conditions to solve for \( X(x) \). After separation,
    we have a second order differential equation \( X \), which yields an
    auxiliary equation whose roots depend on the sign of \( \lambda \). This
    yields a family of solutions \( X_1(x),X_2(x),\dots \) (eigenfunctions)
    corresponding to the \( \lambda \) values \( \lambda_1,\lambda_2,\dots \)
    (eigenvalues).
  \item Find \( T_n(t) \) for each \( \lambda_n \).
  \item Solutions should look like \( u_n(x,t) = X_n(x)T_n(t) \).
  \item Use superposition and Fourier series (if needed) to find the rest of
    the arbitrary constants \( c_n \) that fit the initial conditions.
\end{enumerate}

\subsubsection*{Continuing the Previous Example}
\begin{align*}
  u_t &= \beta u_{xx} \\
  u(0,t) &= u(L,t) = 0 \\
  u(x,0) &= f(x) = 1
\end{align*}
We can apply the initial condition starting from the solution we acquired
earlier.
\begin{align*}
  u(x,t) &= \sum_{n=1}^{\infty}c_n\sin(\frac{n\pi x}{L})
    \e^{-(\frac{n\pi}{L})^2\beta t} \\
  u(x,0) &= \sum_{n=1}^{\infty}c_n\sin(\frac{n\pi x}{L}) = 1 = f(x)
\end{align*}
This becomes a Fourier Sine Series problem with \( L = 1 \) and \( b_n = c_n \).
\begin{align*}
  c_n &= \frac{2}{L}\int_{0}^{L}f(x)\sin(\frac{n\pi x}{L})\diff{x} \\
  &= 2\int_{0}^{1}1\sin(n\pi x)\diff{x} \\
  &= \frac{2}{n\pi}(1-\cos(n\pi))
\end{align*}
We can substitute this back into the formal solution.
\begin{align*}
  u(x,t) &= \sum_{n=1}^{\infty}c_n\sin(\frac{n\pi x}{L})
    \e^{-(\frac{n\pi}{L})^2\beta t} \\
  &= \frac{2}{\pi}\sum_{n=1}^{\infty}\frac{1}{n}(1-\cos(n\pi))
    \sin(n\pi x)\e^{-\beta(n\pi)^2t}
\end{align*}
This is a solution to the boundary value problem when \( f(x) = 1 \) and
\( L = 1 \).

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, please contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
