\documentclass{math}

\usepackage{tikz}

\title{CSCI 251: Concepts of Parallel and Distributed Systems}
\author{Alvin Lin}
\date{December 4th, 2017}

\begin{document}

\maketitle

\section*{Peer to Peer Architectures}
File distribution: client-server versus P2P. For the following models,
\( \mu_s \) is the server upload rate, \( N \) is the number of clients, \( F \)
is the file size, \( d_min \) is the minimum client download rate, and
\( \mu_i \) is the upload rate of the i'th client.

\subsection*{Client-Server Architecture}
\begin{itemize}
  \item Server transmission: the server must sequentially send (upload) \( N \)
  file copies. Time to send one copy: \( \frac{F}{\mu_s} \). Time to send
  \( N \) copies: \( \frac{NF}{\mu_s} \)
  \item Client: each client must download the file copy. If the minimum client
  download rate is \( d_{min} \), then the minimum client download time is
  \( \frac{F}{d_{min}} \).
\end{itemize}
The total time to distribute is lower bounded by the maximum of
\( \frac{NF}{\mu_s} \) and \( \frac{F}{d_{min}} \).

\subsection*{Peer to peer architecture}
\begin{itemize}
  \item Server transmission: the server must upload at least one copy. This
  requires \( \frac{F}{\mu_s} \) time.
  \item Client: each client must download a file copy, taking a minimum of
  \( \frac{F}{d_{min}} \). As an aggregate, they must download \( NF \) bits
  with the max upload rate being \( \mu_s+\sum\mu_i \).
\end{itemize}
The total time to distribute is lower bounded by the maximum of
\( \frac{F}{\mu_s} \), \( \frac{F}{d_{min}} \), and
\( \frac{NF}{\mu_s+\sum\mu_i} \). Compared to the client-server model, whose
total time increases linearly with the number of clients, the total
distribution time for peer to peer architectures increases logarithmically.

\subsubsection*{BitTorrent}
In BitTorrent's file distribution system, the file is divided into 256 kilobyte
chunks. Peers in torrents send and receive file chunks. A \textbf{tracker}
tracks peers participating in the \textbf{torrent}, the group of peers
exchanging chunks of a file. A peer joining the torrent starts with no chunks,
but will accumulate them over time from other peers. It first registers with
the tracker to get a list of peers and connects to a subset of peers. While
downloading, the peer uploads chunks to other peers. The peer may change peers
with whom it exchanges chunks, and once it has the entire file, it may leave
or remain in the torrent to help propagate chunks.

\section*{Reminders}
Project 2 is due Friday, December 8th.

\noindent Professor Mohan Kumar: \\
\url{mjkvcs@rit.edu} \\
\url{https://cs.rit.edu/~mjk} \\

\noindent Rahul Dashora (TA): \\
\url{rd5476@mail.rit.edu} \\

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, please contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
