\documentclass{math}

\title{CSCI 251: Concepts of Parallel and Distributed Systems}
\author{Alvin Lin}
\date{October 25th, 2017}

\begin{document}

\maketitle

\section*{Global States}
A cut is nothing more than a snapshot of a distributed system. For a cut to be
consistent, the events and messages must be a in state such that there is no
state that depends on a state not inside the cut.

\subsection*{Monitor Process}
Assumtions:
\begin{itemize}
  \item During the snapshot operation, there are no faults.
  \item There is a finite time for messages to flow from one process to another.
  \item Any process can start the monitoring process.
\end{itemize}

\subsection*{Chandy \& Lampon Algorithm for Monitoring}
Marker Process (initiating process):
\begin{enumerate}
  \item Marker message is sent by the initiating process.
  \item The marker process records its state and sends a marker message
  \( m \) on every outgoing channel before it sends to any other process.
\end{enumerate}
Receiving Process \( p_j \):
\begin{enumerate}
  \item If \( p_j \) has not recorded its state, \( p_j \) records its state
  now and starts recording messages arriving on all the other incoming
  channels.
  \item If \( p_j \) has already recorded its state, it records the state of
  each channel (recording of messages received since it saved its state).
\end{enumerate}

\subsection*{Coordination and Agreement}
\begin{itemize}
  \item Mutual Exclusion
  \item Multicasting
  \item Consensus
  \item Performance: Latency, bandwidth consumption, energy consumption
\end{itemize}
With a mutual exclusion, only one process has access to a resource at any given
time.

\section*{Reminders}
Check MyCourses for details on Project 2. \\
\noindent Professor Mohan Kumar: \\
\url{mjkvcs@rit.edu} \\
\url{https://cs.rit.edu/~mjk} \\

\noindent Rahul Dashora (TA): \\
\url{rd5476@mail.rit.edu} \\

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, please contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
