\documentclass[letterpaper, 12pt]{math}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}

\title{Sets}
\author{Alvin Lin}
\date{Discrete Math for Computing: January 2017 - May 2017}

\begin{document}

\maketitle

\section*{Sets}
\textbf{Sets} are fundamental discrete structures. A set is an unordered
collection of objects called \textbf{elements} of the set. A set is said
to contain its elements. We write \( a \in A \) if \( a \) is an element of the
set \( A \). We write \( a \notin A \) if \( a \) is not an element of the set
\( a \). Sets are usually given in uppercase letters while elements are
generally given in lowercase letters. \par
There are several ways to describe sets:
\begin{enumerate}
  \item Roster method
  \item Set builder notation
  \item Venn diagrams
\end{enumerate}

\subsection*{Roster Notation}
The set of all vowels is:
\[ V = \bigg\{a,e,i,o,u\bigg\} \]
The set of all odd positive integers less than 12 is:
\[ O = \bigg\{1,3,5,7,9,11\bigg\} \]
Other examples:
\[ U = \bigg\{dog,cat,fish\bigg\} \]
\[ dog \in U \]
\[ chicken \notin U \]

\subsection*{Set Builder Notation}
\[ O = \bigg\{x|x\ is\ an\ odd\ positive\ integer\ less\ than\ 100\bigg\} \]
\begin{align*}
  [a,b] &= \bigg\{x|a \leq x \leq b \bigg\} \\
  (a,b] &= \bigg\{x|a < x \leq b \bigg\} \\
  [a,b) &= \bigg\{x|a \leq x < b \bigg\} \\
  (a,b) &= \bigg\{x|a < x < b \bigg\}
\end{align*}

\subsection*{Set Equality}
Two sets are equal if and only if they contain the same elements. You must show
that the two sets are subsets of each other. Thus \( A \) and \( B \) are equal
if:
\[ \forall{x}(x \in A \leftrightarrow x \in B) \]
or
\[ A \subseteq B \wedge B \subseteq A \]
We write \( A = B \) when sets are equal.
\[ \bigg\{1,2,3\bigg\} = \bigg\{3,1,2\bigg\} \]
\[ \bigg\{1,2,3\bigg\} \neq \bigg\{1,2,4\bigg\} \]

\subsection*{Venn Diagrams}
Venn Diagrams are useful for showing relations between sets. The universal
set \( U \) contains all elements. The set \( V \) contains the set of all
vowels.
\begin{center}
  \begin{tikzpicture}
    \draw (0,0) -- (4,0) -- (4,4) -- (0,4) -- (0,0);
    \draw (2,2) circle (1.5cm);
    \node at (2,2) {a,e,i,o,u};
    \node at (0.25,3.75) {\( U \)};
  \end{tikzpicture}
\end{center}

\subsection*{The Empty Set}
The empty set \( \varnothing \) is the set consisting of no elements (null set).
\par
\textbf{Theorem:} If \( S \) is a set, then \( \varnothing \subseteq S \) and
\( S \subseteq S \).

\subsection*{Subsets}
The set \( A \) is a subset of the set \( B \) if and only if every element of
\( A \) is also in \( B \), notated as \( A \subseteq B \)
\[ \forall{x}(x \in A \to x \in B) \]
\[ \bigg\{1,2,3\bigg\} \subseteq \bigg\{1,2,3,4\bigg\} \]
\[ \bigg\{1,2,4\bigg\} \subseteq \bigg\{1,2,5\bigg\} \]
Sets may have other sets as members:
\[ A = \{\varnothing, \{a\}, \{b\}, \{a,b\}\} \]
\[ a \notin A, b \notin A \]
\[ \{a\} \in A, \{b\} \in A \]

\subsection*{The Size of a Set}
Let \( S \) be a set. If there are exactly \( n \) distinct elements of \( S \),
then we say that the \underline{size} of \( S \) is \( n \). Written as
\( |S| = n \). It is also called the \textbf{cardinality} of the set. In this
case, \( S \) is finite. \par
Let \( A \) be the set of letters in the alphabet. \( |A| = 26 \). \par
A set is said to be infinite if it is not finite. For example, the set of
integers is infinite.

\subsection*{Power Sets}
Given a set \( S \), the power set of \( S \) is the set of all subsets of
\( S \). This is denoted by P(S).
\[ P(\{0,1,2\}) = \{ \varnothing, \{0\}, \{1\}, \{2\}, \{1,2\}, \{1,3\}, \{2,3\},
   \{1,2,3\}\} \]
For a set \( S \) with \( |S| = n \), \( |P(S)| = 2^{n} \).

\subsection*{Cartesian Product}
Let \( A \) and \( B \) be sets. The \textbf{Cartesian Product} of \( A \) and
\( B \), denoted \( A \times B \), is the set of all \underline{ordered} pairs
\( (a, b) \) where \( a \in A \) and \( b \in B \).
\[ A \times B = \{(a,b)\ |\ a \in A \wedge b \in B\} \]
Let \( A = \{1,2\} \quad B = \{a,b,c\} \):
\[ A \times B = \{(1,a),(1,b),(1,c),(2,a),(2,b),(2,c)\} \]
\[ B \times A = \{(a,1),(a,2),(b,1),(b,2),(c,1),(c,2)\} \]
\( A \times B \neq B \times A \) since the Cartesian Product is ordered.
\( (1,a) \in A \times B \) but \( (1,a) \notin B \times A \).
Also note that the cardinality of a Cartesian Product is as such:
\[ |A \times B| = |A||B| \]

\begin{center}
  If you have any questions, comments, or concerns, please contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
