\documentclass{math}

\title{Introduction to Cryptography}
\author{Alvin Lin}
\date{January 2018 - May 2018}

\begin{document}

\maketitle

\section*{Cryptography}
Desired security properties in the digital world:
\begin{itemize}
  \item confidentiality, secrecy
  \item data integrity
  \item authentication, of data origin and entity
  \item non-repudiation
\end{itemize}
Cryptography is an important, but only a relatively small part of security:
\begin{itemize}
  \item the right choice of tools is hard
  \item implementation errors are common
  \item a variety of side-channel attacks can bypass the best cryptography
  \item social engineering
\end{itemize}

\subsection*{Unkeyed, Symmetric-Key, and Public-Key}
Primitives, algorithms, and protocols can be \textbf{unkeyed},
\textbf{symmetric-key}, or \textbf{public-key}.

\subsubsection*{Unkeyed}
This includes hashing and the SHA-family of algorithms. It can be used for
signing and the generation of random sequences.

\subsubsection*{Symmetric Keys}
All encryption schemes from ancient times until 1976 were symmetric ones. The
same key is used for both encryption and decryption. Encryption and decryption
are inverse operations if the same key is used on both sides. Keys must be
transmitted via some other secure channel.
\begin{itemize}
  \item Block ciphers since the 1970s: IBM's Lucifer, DES (Data Encryption
  Standard), IDEA (International Data Encryption Standard), AES (Advanced
  Encryption Standard).
  \item Stream ciphers: RC4, also can come from counter mode of block ciphers
  or hash functions.
  \item MAC, HMAC: message authentication codes.
  \item PRNG: pseudorandom number generators
\end{itemize}

\subsubsection*{Public Key (Asymmetric)}
\begin{itemize}
  \item Public-key cryptosystems: RSA (Rivest, Shamir, Adleman), ElGamal,
  McEliece cryptosystems, ECC (elliptic curve cryptosystems).
  \item Signatures: DSS/DSA (Digital Signature Standard/Algorithm), ECDSA
  (Elliptic Curve Digital Signature Algorithm).
  \item PKI: public key infrastructure, DH (Diffie Hellman key agreement),
  key management, and distribution.
  \item Homomorphic cryptography: Paillier, Gentry
\end{itemize}

\subsection*{Main Public-Key Systems in Use}
RSA by Rivest-Shamir-Adleman (1977) has an edge of ECC because:
\begin{itemize}
  \item it is simple and well understood
  \item links nicely to basic number theory
  \item deployed earlier on many systems
\end{itemize}
ECC by Koblitz-Miller (1985) has an edge over RSA because:
\begin{itemize}
  \item it uses short keys (163+ bits vs 1024+ bits for RSA)
  \item delivers much better performance
  \item uses great theory of elliptic curves on top of classical number theory
  used by RSA
\end{itemize}

\subsection*{Other composite and special functionalities}
\begin{itemize}
  \item Zero-knowledge protocols
  \item Authenticated encryption CAESAR competition
  \item Electronic cash: untraceable, no double-spending, bank-shop-customer
  roles, central banks
  \item Cryptocurrencies: Bitcoin, Zerocoin, Litecoin, Darkcoin, etc
  \item Electronic voting: no individual vote auditing
  \item Oblivious transfer: two millionaires problem
  \item Quantum and post-quantum cryptography, quantum key distribution,
  quantum computing
\end{itemize}
The majority of today's protocols are hybrid schemes. We use symmetric ciphers
for encryption and message authentication, and asymmetric ciphers for key
exchange and digital signatures.

\subsection*{Why do we need Cryptanalysis?}
There is no mathematical proof of security for any practical cipher. The only
way to have assurance that a cipher is secure is to try to break it (and fail).
\textbf{Kerckhoff's Principle} is paramount in modern cryptography: A
cryptosystem should be secure even if the attacker knows all details about the
system, with the exception of the secret key. In order to use Kerckhoff's
Principle in practice, only use widely known ciphers that have been
cryptanalyzed for several years by good cryptographers.
\par It is tempting to assume that a cipher is ``more secure'' if its details
are kept secret. However, history has shown time and again that secret ciphers
can almost always be broken once they have been reverse engineered.

\subsection*{Cryptanalysis: Attacking Cryptosystems}
Classical attacks involve mathematical analysis or brute-force attacks.
Implementation attacks try to extract the key through reverse engineering
and various side-channel attacks. Social engineering attacks try to trick
users into giving up passwords and authentication details.

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, please contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
