\documentclass{math}

\title{Introduction to Computer Vision}
\author{Alvin Lin}
\date{August 2018 - December 2018}

\begin{document}

\maketitle

\section*{Local Features}
We want to find patches that are ``worth representing'' to match from image to
image. These will represent either textures or objects. Ideally, they should be
covariant to translation, rotation, and scale. If the image is translated,
rotated, or scaled, the neighborhoods still preserve their unique properties.
They should also be localizable in translation, rotate, and scale, meaning we
can estimate the position, orientation and size of the patch.

\subsection*{Good Features}
\begin{itemize}
  \item Repeatability: the same feature can be found in several images despite
    geometric and photometric transformations.
  \item Saliency: each feature has a distinctive description.
  \item Compactness and efficiency: many fewer features than image pixels.
  \item Locality: a feature occupies a relatively small area of the image and is
    robust to clutter and occlusion.
\end{itemize}
Feature points are often used for image alignment, 3D reconstruction, motion
tracking, robot navigation, indexing and database retrieval, and object
recognition.

\subsection*{Finding Corners}
A key property of corners is that in the region around a corner, the image
gradient has two or more dominant directions. A strategy to find corners is to
find centers and estimate the scale from the center. Once we have that, we can
estimate the orientation. In a small enough window, a corner has a large
gradient change in all directions. Consider shifting a small window \( W \) by
\( (u,v) \):
\[ E(u,v) = \sum_{(x,y)\in W}[I(x+u,y+v)-I(x,y)]^2 \]
This yields an error comparing each pixel before and after. If we assume a
small amount of motion we can use a first order Taylor Series expansion of
\( I \):
\begin{align*}
  I(x+u,y+v) &\approx I(x,y)+\pdiff{I}{x}u+\pdiff{I}{y}v \\
  &\approx I(x,y)+[I_x\ I_y]\begin{bmatrix}u \\ b\end{bmatrix} \\
  E(u,v) &\approx \sum_{(x,y)\in W}\left[I(x,y)+[I_x\ I_y]
    \begin{bmatrix}u \\ b\end{bmatrix}-I(x,y)\right]^2 \\
  &\approx \sum_{(x,y)\in W}\left[[I_x\ I_y]\begin{bmatrix}u \\ v\end{bmatrix}\right]^2 \\
  &= \sum_{(x,y)\in W}[u\ v]\begin{bmatrix}
    (I_x)^2 & I_xI_y \\
  \end{bmatrix}\begin{bmatrix}u \\ v\end{bmatrix}
    I_yI_x & (I_y)^2
\end{align*}

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, please contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
