\documentclass{article}

\title{Bullshit}
\author{Alvin Lin}
\date{Introduction to Philosophy: January 2017 - May 2017}

\begin{document}

\maketitle

\subsubsection*{Identify some instance of bullshit and write 1 page on why it
is bullshit.}

\textbf{Premise:} Religion, as a whole, is bullshit. For the purposes of this
argument, I will define religion as a ``belief in supernatural entities or
forces with some effect on the natural world''.

In class, we discussed how a person could be bullshitting regardless of the
truth of his/her words. The amount of bullshit in a person's words is irrelevant
to the truth of his statement, but rather their relation to those words, their
relation to their audience, and their intent in their delivery. A bullshitter
by definition seeks to promote some agenda through a manipulation of the
way they are perceived. \par
There is no better example for this than religion itself. Over the past few
millenia, religion has emerged as a dominant force in society, imposing rules
and guidelines on a person's behavior and mindset. Religion came about as a
Stone Age method of social structure and control, and thus its agenda has
always been establishing social control. \par
Like bullshit, religion is wholly unconcerned with the facts, or lack
thereof, in their dogma. With the exception of a small portion of the devout,
few religious people believe in the literal truth of their religious dogma.
Religious scripture very easily falls apart under any scrutiny. \par
It is trivial to establish that religion is unconcerned with truth, what about
morality? Many profess that religious doctrines and texts are meant to be read
metaphorically for the purpose of moral guidance. Well, like bullshit, even if
religion professes to be an upstanding source of metaphorical objective
morality, it still has an agenda unrelated to what its moral strictures are
conveying. As in the case of the monotheistic Abrahamic religions, one can very
easily find tribalistic elements of the ``us vs. them'', ``my god is the only
god'' mentality, which is fueled by the religion's need to self-perpetuate and
claim to legitimacy. \par
Religion creates a virtue out of faith, and like bullshit, requires a quality
of self-deception where one must shut out contradictory information for the
purpose of the religion.

\end{document}
