\documentclass{math}

\usepackage{tikz}

\title{University Physics 1A}
\author{Alvin Lin}
\date{September 22nd, 2017}

\begin{document}

\maketitle

\section*{Modified Atwood Machine}
Two masses (cart \( m_A \), hanging block \( m_B \)) are connected by an ideal
string passing over an ideal pulley, as shown. Note that an ``ideal string'' has
the same tension force \( T \) everywhere along its length, but that force pulls
to the right on sliding cart \( m_A \) and up on the hanging mass \( m_B \).
Assume that the friction on the horizontal surface is negligible. Figure
out the forces acting on the cart. Fill in the table.
\begin{center}
  \begin{tabular}{|c|c|c|}
    \hline
    \textbf{Name of force on cart} & \textbf{x-component} &
      \textbf{y-component} \\ \hline
    gravity & 0 & \( -m_Ag \) \\ \hline
    normal & 0 & \( N \) \\ \hline
    tension & \( T \) & 0 \\ \hline
    total & \( T \) & \( N-m_Ag \) \\ \hline
  \end{tabular}
\end{center}
Now do the same for the forces acting on the weight.
\begin{center}
  \begin{tabular}{|c|c|c|}
    \hline
    \textbf{Name of force on weight} & \textbf{x-component} &
      \textbf{y-component} \\ \hline
    gravity & 0 & \( -m_Bg \) \\ \hline
    tension & 0 & \( T \) \\ \hline
    total & 0 & \( T-m_Bg \) \\ \hline
  \end{tabular}
\end{center}
Apply Newton's Second Law. You should end up with a set of two useful
equations.
\[ T = m_Aa_A \]
\[ T-m_Bg = m_Ba_B \]
Combine the equations to eliminate variables:
\begin{align*}
  T &= m_Aa_A \\
  T-m_Bg &= m_Ba_B \\
  m_Aa-m_Bg &= -m_Ba \\
  m_Aa+m_Ba &= m_Bg \\
  a(m_A+m_B) &= m_Bg \\
  a &= \frac{m_B}{m_A+m_B}g
\end{align*}

\section*{Friction}
Kinetic Friction:
\[ F_{kinetic} = \mu_kN \]
\( \mu_k \) is the coefficient of kinetic friction. \\
Static friction is as large as it needs to be to stop motion.
\[ F_{static} = \mu_sN \]

\section*{Reminders and Homework}
Complete the homework on TheExpertTA and WebAssign. \\
\textbf{Remember to bring the Activities Manual}

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, please contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
