\documentclass[letterpaper, 12pt]{article}

\usepackage{hyperref}
\usepackage{ulem}

\title{ISTE 110: Ethics in Computing}
\author{Alvin Lin}
\date{January 30th, 2017}

\begin{document}

\maketitle

\subsection*{Ethics in Information Technology}
Most computing companies and social networking companies have a code of ethics
in the form of a privacy policy. The concept of ethics in society has been
in the news recently for many reasons. We are organized under principles and
values for the common goal of keeping society safe. Within that bigger picture,
each individual must make ethical decisions. \par
``Ethics and ethical principles extend to all spheres of human activity. They
apply to our dealings with each other, with animals anad the environment. They
should govern our interactions not only in conducting research but also in
commerce and politics.''

\subsection*{What is ethics?}
Ethics is more than examining what is right versus what is wrong. A focal point
is ``the good life'' and the right conduct in an idealistic society.

\subsection*{Ethics versus Morals}
\textbf{Ethics} are the principles and values a society deems as important to
living ``the right way'' (societal perception). \\
\textbf{Morals} are the rules we create that support ethical standards.
(Personal perception)

\subsection*{Law and Order}
In addition to ethics and morals, we have law and order applied to maintain
those ethical standards. We are concerned with the intent, impact, and outcomes
of ethical behavior.

\subsection*{Why is this important?}
\begin{itemize}
  \item Look at today's culture
  \item How we use technology
  \item Meshing of values at work, at home, and how we use technology
  \item How does this reconcile across generations?
\end{itemize}
Almost everything in our everyday lives is impacted by technology. Different
age groups in our generations grew up with different values, ideals, and
exposure to technology.

\subsection*{Group Presentation Topics}
Choose an information technology topic to present on in groups. Do not do the
following:
\begin{itemize}
  \item The Digital Divide
  \item The First and Fourth amendment
  \item Digital Privacy
  \item Cyberbullying
\end{itemize}

\subsection*{Brainstorm Topics}
\begin{itemize}
  \item \sout{Parental Monitoring}
  \item \sout{The Right to be Forgotten}
  \item \sout{Artificial Intelligence}
  \item \sout{Automation of Jobs}
  \item \sout{Internet Censorship}
  \item \sout{Whistleblowers}
  \item \sout{Cryptocurrency}
  \item \sout{The Internet: Humanity's Greatest Experiment in Anarchy}
  \item \sout{Open Source Code Ownership/Attributions}
  \item \sout{Geographically Restricted Content}
  \item Life Prioritization Routines in Autonomous Vehicles
  \item Combat Drones/Robots
  \item Social Media and its Interaction with Society
  \item NSA Monitoring
  \item Expansion of Genetic Technology
\end{itemize}

\subsection*{Reminders and Homework}
Remember to email catherine.beaton@rit.edu with the following:
\begin{verbatim}
Subject: ISTE 110 Team
Body:
Team Name
Team Members
Team Lead
\end{verbatim}
Come with ideas next class for team contracts.

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, feel free to contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
