\documentclass[letterpaper, 12pt]{article}

\usepackage{hyperref}

\title{ISTE 110: Ethics in Computing}
\author{Alvin Lin}
\date{April 3rd, 2017}

\begin{document}

\maketitle

\section*{Women In Computing}
Lorne Grabher had his last name on a vanity license plate, which was deemed
socially unacceptable and revoked. \\
\href{http://dailycaller.com/2017/03/25/grabher-license-plate-called-socially-unacceptable-in-nova-scotia/}{Grabher License Plate Called Socially Unacceptable In Nova Scotia} \\
Why are women less included in politics?
Nintendo first chose to market in the boy's aisle of stores for their product.
Many people develop an interest in computer scientists through gaming.
Women are also an ostracizing group in games: \\
\href{https://www.lifewire.com/xbox-live-survival-for-female-gamers-3562487}
{Xbox Live Survival for Female Gamers} \\

\subsection*{[Pizza Break]}
Discuss with your groups why such a disparity exists in tech. \\
One of the biggest reasons behind this disparity is culture, particularly
online culture. Historically, boys have gone into science while girls study
``soft'' areas like English. \par
The workplace faces similar problems, which extends to demographics such as
people with disabilities. Also note the recent government meetings regarding
Planned Parenthood, which consisted of almost entirely men.

\subsection*{You can't be it if you can't see it}
There are few women in computing, and women are discouraged because they don't
see other women succeeding in the field. At RIT, there is a 90-10 male-female
disparity. Overall at RIT, it is a little better at 65-35.

\subsection*{Reddit}
Reddit's two pillars are anonymity and free speech. While some parts are very
informative, other parts are toxic and misogynist. A women noted in an
experiment that after posting the same comic with both a male and female
username, she was bashed under the female username and encouraged using the
male username. \par
There are subreddits designed to bring people with similar interests together.
After Anderson Cooper did a piece on subreddit /r/jailbait, which featured
pictures of often underage girls, the public response was explosive and Reddit
actually took down the subreddit.

\section*{Reminders and Homework}
The Lessig Assignment is due Monday, April 10th at 11:59PM.
Work on your final paper and projects.

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, feel free to contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
