\documentclass[letterpaper, 12pt]{article}

\usepackage{hyperref}

\title{ISTE 110: Ethics in Computing}
\author{Alvin Lin}
\date{February 20th, 2017}

\begin{document}

\maketitle

\section*{Think, Pair, Share: Generation of Topic and Paper Outline}
Socrates is renown for his ability to engage people in deep-thought
discussions. He asks a series of questions on a focal point, and assists in
developing critical thinking among participants. The Socratic method can bring
forward thought paths that you may not have considered, or discarded without
recognizing their worth. \par
Your task today is to spend time choosing a topic for your final paper. Do some
research, surf the web, achieve an epiphany. When you have your topic, you can
begin to do a point form outline. Not full sentences. Just thoughts. \par
Upon completing the outline, you now need to find a person to pair with, and
share ideas. It can NOT be a person you sit next to, nor a member of your
team. Branch out. Make new friends. \par

\section*{Scholarly Style Research Paper Outline}
Your content for today's class is posted to MyCourses:
Scholarly Style Research Paper Outline Template (Kibin.com): \\
\url{https://www.kibin.com/essay-writing-blog/research-paper-outline} \par
This is a graded assignment valued at 5 points. You are to complete a research
paper outline, and then share it with a peer (but not a team-mate).
Your time line is as follows: From 5 pm until 5:45 pm, surf, brainstorm, give
deep thought to something that interests you; something you could write about
that makes you enthusiastic. Picking an ``easy'' topic doesn't play out well,
because I can sense the boredom in your writing. At 5:45, find a friend and
have them discuss, in a Socratic method, your thoughts. The other person's goal
is to get you to think on a deeper level than we did on the presentation
outline. \par
I will need one person from the duo to email me with the subject line: ``Essay
Outline'' and put your names in the body of the message. \par
It is possible this will not remain your topic. You'll receive a grade for
whatever outline is placed in the drop box on March 6th. Wednesday, we'll be
going into more detail about the paper, but a starting place for all these
activities begins with practice.

\section*{Basic Research Paper Outline Template}
\begin{enumerate}
  \item Introduction
  \begin{enumerate}
    \item Hook
    \item Research Question
    \item Thesis Statement
  \end{enumerate}
  \item Body Paragraphs
  \begin{enumerate}
    \item Context and History of Research Topic
    \item Existing Arguments
    \begin{enumerate}
      \item First Argument
      \begin{enumerate}
        \item Citation
        \item Citation
      \end{enumerate}
      \item Second Argument
      \begin{enumerate}
        \item Citation
        \item Citation
      \end{enumerate}
      \item Etc.
    \end{enumerate}
    \item Your Argument
    \begin{enumerate}
      \item First Argument
      \begin{enumerate}
        \item Citation
        \item Citation
      \end{enumerate}
      \item Second Argument
      \begin{enumerate}
        \item Citation
        \item Citation
      \end{enumerate}
      \item Etc.
    \end{enumerate}
  \end{enumerate}
  \item Conclusion
  \begin{enumerate}
    \item Summary of Main Points
    \begin{enumerate}
      \item Point 1
      \item Point 2
      \item Etc.
    \end{enumerate}
  \end{enumerate}
\end{enumerate}

\section*{Research Paper on Autonomous Cars and Car Ownership}
\begin{enumerate}
  \item Introduction
  \begin{enumerate}
    \item Autonomous cars have been in the mind of researchers and engineers
      since the mid-20th century. The commercialization of autonomous cars will
      make roads safer and streamline inner-city transportation, but may also
      revolutionize the idea of car ownership in the future.
    \item How will autonomous vehicle fleets affect the concept of car ownership
      in the future?
    \item In the next century, car ownership will decrease as the
      commercialization of autonomous cars will allow for shared ownership of
      a car and on-demand transportation at the click of a button.
  \end{enumerate}
  \item Ethical Theory: Utilitarianism/Rawlsianism
  \begin{enumerate}
    \item Services like Uber, Airbnb, Lyft, and Sidecar are part of a new idea
      of a ``sharing economy''.
    \item \href{https://www.theguardian.com/technology/2016/oct/17/sharing-economy-capitalism-uber-airbnb-ownership}{\underline{The Guardian}}
    \item \href{http://knowledge.insead.edu/responsibility/whos-responsible-the-ethics-of-the-sharing-economy-5034}{\underline{Knowledge Insead}}
    \item \href{https://kenan.ethics.duke.edu/wp-content/uploads/2012/08/Sharing-EconomyTN2015.pdf}{\underline{The Kenan Institute for Ethics}}
    \item \href{http://www.adweek.com/digital/sharing-economy-steeped-ethical-controversy/}{\underline{Adweek}}
    \item \href{http://spectrum.ieee.org/cars-that-think/transportation/self-driving/people-want-driverless-cars-with-utilitarian-ethics-unless-theyre-a-passenger}{\underline{Spectrum IEEE}}
  \end{enumerate}
  \item Body Paragraphs
  \begin{enumerate}
    \item Context: Carnegie Mellon has been experimenting with autonomous and
      semi-autonomous cars since the 1980s. Now, companies like Tesla have made
      cars nearly completely self-drivable using advanced arrays of sensors and
      processors. Autonomous cars can now be connected to the internet and sent
      to locations via an app.
    \item Existing Arguments
    \begin{enumerate}
      \item Already, companies like Uber are challenging the concepts of car
        ownership by providing readily accessible transportation at the click
        of a button.
      \begin{enumerate}
        \item \href{http://www.businessinsider.com/ubers-plans-to-be-cheaper-than-owning-a-car-2015-2}{\underline{Business Insider}}
      \end{enumerate}
      \item Combining this idea with fleets of autonomous cars would only
        accelerate this social change.
      \begin{enumerate}
        \item \href{http://www.businessinsider.com/former-tesla-vp-georg-bauer-autonomy-will-kill-car-ownership-2016-11}{\underline{Business Insider}}
      \end{enumerate}
    \end{enumerate}
    \item Your Argument
    \begin{enumerate}
      \item Autonomous cars will be widely integrated into society in the near
        future anyway simply due to the fact that they are safer and will reduce
        accidents by a significant percentage.
      \begin{enumerate}
        \item \href{https://www.youtube.com/watch?v=WBjY3QGNdAw}{\underline{Veritasium}}
        \item \href{https://www.washingtonpost.com/news/innovations/wp/2017/02/20/the-big-moral-dilemma-facing-self-driving-cars/?utm_term=.802ae11b67c4}{\underline{Washington Post}}
      \end{enumerate}
      \item Fleets of autonomous cars and shared car ownership will prove to be
        cheaper and more convenient than owning a car. An on-demand service
        model will make transportation easier to access and phase out issues
        such as parking space.
      \begin{enumerate}
        \item \href{http://www.businessinsider.com/why-no-one-will-own-a-car-in-25-years-2015-6}{\underline{Business Insider}}
        \item \href{http://www.slate.com/articles/technology/future_tense/2016/10/self_driving_cars_effects_on_cities_depend_on_who_owns_them.html}{\underline{Slate}}
      \end{enumerate}
    \end{enumerate}
  \end{enumerate}
  \item The integration of autonomous cars into a city's infrastructure will
    revolutionize concepts of ownership regarding cars. Transportation can
    eventually be turned into a public service that is cheaper and more
    accessible than personal ownership of a car.
  \begin{enumerate}
    \item Summary of Main Points
    \begin{enumerate}
      \item Companies like Uber are already changing the idea of car ownership.
      \item Autonomous cars will be adopted anyway because they decrease road
        accidents and are many times safer.
      \item Fleets of autonomous cars providing an on demand service model will
        make transportation more accessible and cheaper, solving problems like
        parking spaces.
    \end{enumerate}
  \end{enumerate}
\end{enumerate}

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, feel free to contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
