\documentclass{article}

\usepackage{hyperref}

\title{ISTE 110: Ethics in Computing}
\author{Alvin Lin}
\date{March 6th, 2017}

\begin{document}

\maketitle

\subsection*{The News in Real-Time}
The News in Real-Time is a really big deal today. We are accessing information
in such a fast and brief way that we only receive a splash of the event.
A long time ago, we had to wait for news via newspapers, and it was usually
a day late. Instant gratification is something we take for granted in news.

\subsection*{Good? Bad? Just plain ugly?}
Is our instant gratification fed appropriately? Should it be? Have we become
``media whores''? \par
Journalist Daniel Pearl was kidnapped by Pakistani terrorists and beheaded. The
footage of the beheading was released to the public.

\subsection*{Technology and Trust}
We Shall Never Forget 9/11 - The Kids Book of Freedom - A Graphic Coloring Book
This children's coloring book provides images of reports of 9/11 and the US
killing Bin Laden for the purpose of teaching children.

\subsection*{Has our humor changed?}
We print political cartoons in newspapers and distribute them on the web. Some
people find them offensive, but the Internet has exacerbated this phenomenon.

\subsection*{John Stuart Mill}
\begin{itemize}
  \item Where does the authority of society begin?
  \item ``Everyone who receives the protection of society owes a return for the
    benefit, and the fact of living in society renders it indispensible that
    each should be bound to observe a certain line of conduct towards the
    rest.''
  \item ``The acts of an individual may be hurtful to others or wanting in due
    consideration for their welfare, without going to the length of violating
    any of their constitutional rights. The offender may then be justly
    punished by opinion, though not by law.''
  \item ``As soon as any part of a person's conduct affects prejudicially the
    interests of others, society has jurisdiction over it, and the question
    whether the general welfare will or will not be promoted by interfering
    with it becomes open to discussion.''
  \item ``In all cases, there should be perfect freedom, legal and social, to
    do the action and stand the consequences.''
  \item John Stuart Mill's conception of liberty involved freedom of people
    instead of state control.
  \item Utilitarianism - Greatest Happiness Principle
\end{itemize}

\subsection*{Where is society heading?}
\begin{itemize}
  \item ``A civilization that can thus succumb to its vanquished enemy must
    first have become so degenerate that neither its appointed priests and
    teachers, not anybody else, has the capacity, or will take the trouble, to
    stand up for it.''
  \item Is THIS civilization?
  \item Where are we going in 2020?
  \item The transparency of people and organizations will increase, but that
    will not necessarily yield more personal integrity, social tolerance, or
    forgiveness.
  \item What has the Internet given us? What has it taken from us?
  \item Who is ethically responsible for the release of information without
    fact-checking?
  \item Does the internet facilitate revolution?
\end{itemize}

\subsection*{Reminders and Homework}
Remember to submit your final paper outline on MyCourses.

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, please contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
