\documentclass[letterpaper, 12pt]{article}

\usepackage{hyperref}

\title{ISTE 110: Ethics in Computing}
\author{Alvin Lin}
\date{April 3rd, 2017}

\begin{document}

\maketitle

\subsection*{Online Privacy}
Most people are not aware of the fact that most search engines collect your
search information for targeted advertisement. For games like League of
Legends, data like language is monitored and gathered to counter the game's
notoriously toxic community. \\
\href{http://www.nature.com/news/can-a-video-game-company-tame-toxic-behaviour-1.19647}{Can A Video Game Company Tame Toxic Behavior - Nature.com}

\subsection*{Extra Credit}
Design your own choose-your-own-adventure related to ethical behavior in
gaming.

\subsection*{Emotional Contagion}
A social media company noted that direct interactions were not necessary to
evoke positive or negative emotion. By showing positive posts on a feed,
people tended to post more positive things, and the opposite was true.
Emotions can transfer without direct personal interaction.

\subsection*{Lessig}
Watch the Lessig video and respond to it. Discuss the right and wrong sides
of the privacy argument. What is the relevance to moral decision making with
regard to copyright? What is Lessig's interpretation of locked versus unlocked,
and read-only versus read-write? Where are we now?

\subsection*{Group Projects}
Convene with your group to discuss progress on your group project.

\subsection*{Reminders and Homework}
The Lessig Assignment is due Monday, April 10th at 11:59PM.
Work on your final paper and projects.

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, please contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
