\documentclass[letterpaper, 12pt]{article}

\usepackage{hyperref}
\usepackage{ulem}

\title{ISTE 110: Ethics in Computing}
\author{Alvin Lin}
\date{February 13th, 2017}

\begin{document}

\maketitle

\subsection*{Presentations}
Presentations will begin in the middle of March.
What are elements of a good presentation?
\begin{itemize}
  \item One that is prepared well
  \item One that is fluid and seamless
  \item One that is engaged with the audience
  \item One that has useful content and includes visual elements
  \item One where the presenter doesn't read off the slides
  \item The text of the presentation should merely summarize what is said
  \item The presenter should project their voice
\end{itemize}
Tips:
\begin{itemize}
  \item Know your audience.
  \item Know your information. Start with an outline and expand into it.
  \item You will have 30 minutes to present.
  \item In 10 minutes, the audience will remember three things. You need to
    share what you think is relevant to the audience.
  \item The presentation should end with three or four points that you want
    the audience to take away.
  \item Use visual elements to emphasize what you are talking about.
  \item Practice your presentation and know how long it takes to go through
    each part.
  \item Make eye contact with your audience and engage with the audience. Be
    aware of your body language.
  \item Don't mumble, project your voice, keep your voice at an even pitch and
    speak slowly.
  \item You are telling a story, provide a start, middle, and end.
  \item Use 30pt font, visuals, 10 slides maximum.
\end{itemize}

\subsection*{Example Presentation Outline}
Global Warming:
\begin{enumerate}
  \item Nature vs Human Action
  \item Proponents/Opponents
    \begin{itemize}
      \item political divide
      \item age divide (generational gap)
    \end{itemize}
  \item Science vs ``Reasoning''
    \begin{itemize}
      \item historical data
      \item longitudinal studies
      \item geographical changes
    \end{itemize}
  \item Famous Folks
    \begin{itemize}
      \item Neil deGrasse Tyson
      \item Sarah Palin
      \item Bill Nye
      \item Donald Trump
      \item Leonardo de Caprio
      \item Al Gore
    \end{itemize}
  \item Closing Statements
\end{enumerate}
You have 10 minutes to prepare an example presentation on this.

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, feel free to contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
