\documentclass[letterpaper, 12pt]{article}

\usepackage{hyperref}
\usepackage{ulem}

\title{ISTE 110: Ethics in Computing}
\author{Alvin Lin}
\date{February 8th, 2017}

\begin{document}

\maketitle

\section*{Ethics in Information Technology}

\subsection*{Class Instructions}
Write your team leader's email on your group discussion outline and exchange
with another group. At the end of the class, email your feedback to the
designated email. After you finish, exchange again with the next group.

\subsection*{The First Amendment}
Congress shall make no law respecting an establishment of religion, or
prohibiting the free exercise thereof; or abridging the freedom of speech; or
of the press; or the right of the people peaceably to assemble, and to petition
the government for a redress of grievances.

\subsection*{Democracy}
The net has become a democratizing force with many new voices. It facilities
redistribution of information. Information is no longer owned by the ``fourth
estate'' (traditional media). The internet allows people to explore their
curiosity and connect with others, but also allows for scammers, spam, hate
speech, etc. The ability to speak and be read is amplified by the internet.

\subsection*{Freedom of Speech}
Is freedom of speech borderless? We have limitations in press because there is
enforceable accountability. Online, captchas and age restriction warnings
enforce slight but bypassable limitations.

\subsection*{Reno v. ACLU}
\textit{Reno v. ACLU} was a United States Supreme Court case in which all 9
Justices of the Court voted to strike down anti-obscentiy provision of the
Communications Decency Act (DCA), finding that they violated the freedom of
speech provisions of the First Amendment.

\subsection*{Freedom of Choice}
Are we depending on individuals in the government to make moral/immoral
decisions? If so, why them?

\subsection*{Reminders and Homework}
Watch the ``Sneakers'' movie on MyCourses.
Assignemnt: ``Sneakers'' Movie Reaction Paper, due in Dropbox by Sunday,
February 12th, at 11:59pm. You must submit a Word document, please do not
submit a PDF.

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, feel free to contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
