\documentclass{article}

\usepackage{hyperref}

\title{ISTE 110: Ethics in Computing}
\author{Alvin Lin}
\date{April 10th, 2017}

\begin{document}

\maketitle

\subsection*{Women In Computing}
The perception still exists that the computing field is a place for men. Women
don't have the kind of brain for programming and are better off going into
education, nursing, and liberal arts.

\subsection*{Dave Winer, Programmer/Developer}
``Programming is a very modal activity. To be any good at it you have to focus.
And be very patient. I imagine it's a lot like sitting in a blind waiting for a
rabbit to show up so you can grab it and bringit home for dinner. There is
specializstion in our species. It seems pretty clear that programming as it
exists today is a mostly male thing. Which also raises the obvious question
that perhaps we can make it so that it can better use the abilities of the
other half of our species?''

\subsection*{Marie Wilson}
Marie Wilson is the founder and president emeritus of The White House Project
and said women don't seek positions of power because of a lack of role models
who have gone ahead to forge a path. ``You can't be what you can't see.''

\subsection*{The Brogrammer}
``A programmer who breaks the usual expectations of quiet nerdiness and opts
instead for the usual trappings of a frat-boy: popped collars, bad beer, and
calling everybody ``bro''. Despied by everyone, especially other programmers.''
- Urban Dictionary

\subsection*{Pax Dickinson}
Pax Dickinson is the CEO of Business Insider, ``the epitome of someone you'd
like to punch in the face every time he opens his mouth.'' - Professor Beaton \\
Tweets from Dickinson:
\begin{itemize}
  \item ``Tech managers spend as much time worry about how to hire talented
  female developers as they do worry about how to hire a unicorn.''
  \item ``feminism in tech remains the champion topic for my block list. my
  finger is getting tired.''
  \item ``[regarding a transgendered programmer]: so he's a woman, but he still
  has a Y chromosomes? now who doesn't understand how biology works?''
\end{itemize}
Websites like Valleywag and the Washington Post finally caught wind of these
tweets, and he was fired. In an interview with Anil Dash, he seemed to show
little understanding as to why he should change. Later, said ```misogyny' is
`hatred of women'. It is not misogyny to tell a sexist joke, or to fail to take
a woman seriously, or to enjoy boobies.''

\subsection*{TechCrunch 2013}
The kick-off of Tech Crunch in 2013 demonstrated a ``Tit-stare'' app.

\subsection*{Longorrhea}
``Here's a theory - when you've been getting away with your puerile logorrhea
in this hyper-macho brogrammer vacuum for so long, what's there to stop you?''
- Emily Andrews

\subsection*{Elissa Shevinsky}
\begin{itemize}
  \item ``That's it - I'm finished defending sexism in tech.''
  \item ``I wasn't seeing the problems clearly because I'd been part of the
    industry for too long. I also wanted to focus on getting things done
    rather than on feminist-inspired activitsm. So I made the bros-only
    atmosphere work for me.''
  \item ``I had the greatest time drinking scotch at Google I/O with some of
    the best CTOs in the media industry. They treated me like a bro. I didn't
    want to lose those moments.''
  \item ``Despite all of this, I continued to defend the status quo. I wanted
    to drink scotch with my guy friends and build software. I'm done now. I
    didn't want to think about gender issues but the alternative is tit and
    dick jokes at our industry's most respected events.''
\end{itemize}

\subsection*{Cultural Fit}
``We make sure to hire people who are a cultural fit.'' \\
Shanley Kane tries to translate what start-up founders mean when they reject
people based on ``fit''.

\subsection*{How does it end?}
\begin{itemize}
  \item Gina Trapani: ``Brogrammer culture celebrates frat house values,
    youth over experience and men over women. In the war for hiring great
    talent, the companies that embrace this culture rather than reject it will
    lose. That's a good thing.''
  \item Anil Dash: ``There are people making tech who are positive, ambitious,
    thoughtful, inclusive, curious, empathetic, and self-aware. They're going
    to win.''
\end{itemize}

\subsection*{Reminders and Homework}
Work on your final paper and projects. Write the first draft of your final
paper and upload it to MyCourses.

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, please contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
