\documentclass[letterpaper, 12pt]{article}

\usepackage{hyperref}

\title{ISTE 110: Ethics in Computing}
\author{Alvin Lin}
\date{May 1st, 2017}

\begin{document}

\maketitle

\subsection*{Presentations}
Take notes on the final project group presentations.

\subsection*{Text Based Adventure Game}
Main Points:
\begin{itemize}
  \item Machine Ethics
  \item Who is responsible for autonomous drones who kill humans through some
    decision making algorithm?
  \item Can cars testify as humans in court?
  \item What jurisdiction do we have over autonomous cars?
  \item What jurisdiction do we have over sentient non-human machines?
  \item What defines sentience or autonomy?
  \item What is the difference between sentience and autonomy?
  \item Spontaneous emotional reaction
\end{itemize}

\subsection*{Machine Learning and Utilitarianism}
Main Points:
\begin{itemize}
  \item Utilitarianism is a concept that the most moral decision is one that
    provides the greatest benefit for everyone.
  \item Machine learning and computerized decision making may be immoral for
    life altering decisions such as employment, imprisonment, and life or death.
  \item If computers could manage all the variables in our lives, they could
    make decisions for us that generate the greatest amount of happiness.
  \item Neural networks, learning, and backpropagation. You train a neural
    network with input data and target outputs, and the neural network will
    adjust its weights until it can replicate the desired output with some
    accuracy threshold.
  \item \url{https://github.com/JacobPimental/trolleybot}
\end{itemize}

\subsection*{Reminders and Homework}
Work on your final paper.

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, please contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
