\documentclass[letterpaper, 12pt]{article}

\title{Opinion Piece}
\author{Alvin Lin}
\date{January 25th, 2017}

\begin{document}

\maketitle

With the inauguration of Donald Trump concluded, it's almost blindingly obvious that neither his policies nor his appointed department heads are going to be good for the nation. Given the various conflicts of interest among Donald Trump and Rex Tillerson (Secretary of State), the blatant inexperience of Betsy DeVos (Education Secretary), and the fact that Scott Pruitt (head of the EPA) denies climate change, the American people have no guarantee that he or his cabinet are going to take any action to progress the nation. Trump has even backed off on his own campaign promises, and while this is classic demagoguery, it just goes to show that Trump clearly does not have the interests of his supporters, let alone the nation, in mind. \par
The administration's denial of climate change is evidence of the influence that oil companies have over the government. Oil is going to run out within the next century or two, and countless scientific studies have shown that the continued use of resources like coal and oil are damaging the environment. The fact that funding for clean energy has been slashed is indicative of the shortsightedness of the administration. These actions serve to bolster the profits of oil companies at the cost of the environment. Though oil and coal provide jobs and stability in the short term, it shows a ``not my generation, not my problem'' attitude that highlights a selfishness in both the voters and the administration. \par
The argument can be made that this is a product of democracy since people vote in the direction of their own interests, and while oil may provide the short term jobs and stability they want, it will only create problems that the generations after us will have to clean up.

\end{document}
