\documentclass{math}

\title{Programming Language Concepts}
\author{Alvin Lin}
\date{January 2018 - May 2018}

\begin{document}

\maketitle

\subsection*{Introduction}
Wikipedia lists about 300 programming languages, while other sources list
thousands. Joseph Marie Jacquard (18th century) was the first to make a
programmable device (a loom).

\subsection*{Why study programming languages?}
\begin{itemize}
  \item Langauge actually determines what we think, both natural and artificial
  languages.
  \item To study languages it to study how to design and implement languages.
  \item Different languages support different \textit{idioms}.
  \item Learning different idioms gives you more design choices, even when not
  using a language with support for them.
\end{itemize}
What influences and individual or group to use a particular programming
language?
\begin{itemize}
  \item Organizational mandate
  \item Pre-existing components
  \item Developer skillset
  \item Design choices
  \item Implementations, shortcomings found, and dire emergencies
\end{itemize}

\subsection*{Some Application Domains}
\begin{itemize}
  \item Natural sciences and engineering, number crunching: FORTRAN
  \item Business, financial calculations, report creation, databases:
  COBOL, RPG, Excel
  \item Intelligent systems: LISP, Prolog
  \item Systems/Embedded hardware: C
  \item Quick-and-dirty programs: shell, BASIC, Ruby, Python, PHP, JavaScript,
  visual languages
  \item Special purpose: PostScript, GPSS
\end{itemize}

\subsection*{Computer Language Classes}
Imperative:
\begin{itemize}
  \item Machine/Assembly languages
  \item ``Classic'': FORTRAN, Pascal, BASIC, C
  \item Object Oriented: Smalltalk, C++, Eiffel, Java, Python
\end{itemize}
Declarative:
\begin{itemize}
  \item Functional: ML, Haskell
  \item Dataflow: spreadsheets
  \item Logic: constraint-based, Prolog
\end{itemize}

\subsection*{From Source to Execution}
What processes are involved to get a program's source code to execute?
\begin{itemize}
  \item Compilation
  \item Assembly
  \item Linking
  \item Loading
  \item Interpretation
\end{itemize}

\subsection*{Shells vs Traditional Programming Languages}
\begin{itemize}
  \item Shells interpret source code.
  \item They tend to bind variables to names very late.
  \item Variable names and statements can be constructed at run time.
\end{itemize}

\subsection*{Reminders}
\begin{itemize}
  \item Read Scott Text 1.1-1.5
  \item Brush up on Python
  \item Scheme, download Dr Racket. Read Scott text 11.3
\end{itemize}

\begin{center}
  You can find all my notes at \url{http://omgimanerd.tech/notes}. If you have
  any questions, comments, or concerns, please contact me at
  alvin@omgimanerd.tech
\end{center}

\end{document}
